### 2011-02-07 Documentation/atropos/ResponseToReviewers.tex

\noindent
My coauthors and I would like to thank the editors and reviewers for
handling our manuscript entitled ``An Open Source Multivariate Framework for $n$-Tissue
  Segmentation with Evaluation on Public Data.''  The reviews were extremely helpful in highlighting both
the strengths and weaknesses of the previous version of the paper.  In
summary, the review found the work to be worthy of publication, with
revisions that focused on improving clarity of some key points. We make every effort, in this response letter and the
associated revision, to clarify our work and improve the paper
according to the reviewers' suggestions. We believe that we carefully
revised the manuscript to address all concerns and provide responses
to the reviewers' comments, which you will find below.   We
also elected to add detail regarding the ICM strategy
available in Atropos. Thus, we would
like to resubmit the revised manuscript which we hope will be suitable
for publication. \\

\noindent
The revision of the paper includes:
\begin{enumerate}
\item  A new figure revision highlighting ICM and providing a simple
  starter example for Atropos usage. 
\item Typo fixes in the equations. 
\item  Many minor changes in the body of the text.  
\item  The references were altered according to reviewer suggestions.
\end{enumerate}
We express our sincere gratitude to
the editorial staff and to the reviewers for their efforts in
evaluating this paper. \\

\noindent
Sincerely,\\

\noindent
Brian Avants, Ph.D.\\
Penn Image Computing and Science Laboratory \\ 
University of Pennsylvania\\
3600 Market Street, Ste 370\\
Philadelphia, PA 19104
\newpage
\begin{center}
\LARGE{{\bf Response To Reviewers}}
\end{center}
\section{Reviewer 1}
We thank reviewer 1 for his/her careful reading for technical content
and for improving our work.\\
\newline
\newline
{\em 1) Page 5, 2nd line from bottom. What are silver standard
  labellings?}
\newline
\newline
We've changed this line to be more specific in the current revision.
\newline
\newline
Done.  
\newline
\newline
{\em 2) Page 10, Equations 8 - 11. These follow standard
  approaches. However, Equation 11 does not seem consistent with [8]
  and [10]. If delta(x\_i,x\_j)=1 for x\_i=x\_j, then U(x) as defined in
  [10] increases, and p(x) as defined in [8] decreases - hence
  labelling similar regions seems to be penalised. However, things
  seem consistent if equation 11 is changed as in the standard graph
  cuts approaches - e.g. see equation 3 Boykov \& Jolly ICCV 2001.}
\newline
\newline
We appreciate the reviewer catching this error in the equation.  If $x_i = x_j$ then $V=0$ and is
1 otherwise, as is standard in Besag's approach. 
\newline
\newline
{\em 3) Page 10, Equation 11. Computing the value of U(x) only
  considers if neighbours are the same or not. However, as this paper
  deals with segmenting a large number of labels why was an approach
  such as that taken in Noe \& Gee IPMI 2001 not considered -
  i.e. learning the probability of labels being neighbours and using
  these?}
\newline
\newline
This is an interesting point and we intend to investigate this
direction of research in future work.  In fact, some basic
functionality along these lines is available within the guts of the
code but additional comparative research into the myriad partial volume
estimation models that have been proposed is required to determine the
best approach.  Thus, we added comments in the discussion to indicate
our interest in pursuing partial volume models but elect not to pursue
this task for this resubmission. 
\newline
\newline
{\em 4) Page 12, lines 10-14. Authors comment that applying graph cuts
  to more than one label is not globally optimal and that EM is a
  reasonable and efficient alternative. Although this is true,another
  property of graph cuts is that it is fast. How does the EM approach
  compare in terms of speed?}
\newline
\newline
Fast is a relative term and we cannot comment on the speed of a method
for which we do not have an optimal implementation.  
The main computational challenge with graph cuts is memory.  Because one needs to
solve multiple shortest path problems---and store the solution of
each---the potential for memory bloat is high.  Even in EM, memory
concerns emerge when enough labels are used (e.g. in our examples).
In theory, though, EM computations are embarrassingly parallel which is
not a property shared by graph-based methods.  As such, one would
expect even an optimally implemented multi-label graph cut method to
lag in speed behind an optimally implemented EM segmentation.

{\color{blue}{ We would also
like to point out that we have had prior experience with graph cuts (see,
for example, our implementation available at http://hdl.handle.net/1926/1503
and related publications:  Zhuang Song, Nicholas J. Tustison, Brian B. Avants, 
James C. Gee: Integrated Graph Cuts for Brain MRI Segmentation. MICCAI (2) 2006: 831-838
and Zhuang Song, Nicholas J. Tustison, Brian B. Avants, James C. Gee: 
Integrated Graph Cuts for Brain MRI Segmentation. MICCAI (2) 2006: 831-838) and
our preference for our current EM-based approach is certainly not restricted to 
speed considerations.  As enthusiastic supporters of open source software,
patent issues constrain the use of graph cuts where no such limitations exist
for EM-based approaches.}}

@@

extension from univariate to multivariate probability. 
\newline
\newline
{\em 6) Page 17. Penultimate line. ``Three likelihood classes have been
  developed, one parametric and one ...''. Should this be ``Three

@@

{\em 8) I could not verify the validity of all the references as they were not numbered neither were they in alphabetical order. }
\newline
\newline
{\color{blue}{We apologize.  In preparing the original
manuscript, we were unable to find a .bst for the
Neuroinformatics journal so we created our own using
natbib and did not use all the options to facilitate
the reviewing process.  The references are now numbered
in the manuscript and in the bibliography section and
are ordered as they appear in the manuscript.   }}

@@

{\em I have only two concerns with the article. First, there is almost
  no explanation of any of the figures in the body of the text, which
  raises questions about the significance of the findings, such as
  those of Figure 4 (``The results show that the PriorProbabilityMaps

@@

associated with each figure.  {\bf FIXME}
\newline
\newline
{\em Second, why is it that ``In this study, we labeled the brain web (sic) template very sparsely, quickly and crudely and did not expect highly accurate results.''

@@

\newline
\newline
We agree with the reviewer and removed this comment and the evaluation
results.  We added: 
    \begin{quote}
    We provide this capability to allow the user to implement

@@

    an interactive editing and segmentation loop.  The user may
    run Atropos with sparse manual label guidance, evaluate the
    results, update the manual labels and repeat until achieving the desired
    outcome.  This processing loop be performed easily with, e.g., ITK-SNAP.
    \end{quote}
{\em On page 22: ``The advancements introduced with N4 permit such an
  adaptive integration with Atropos.'' What relevant advances are

@@

  introduced with respect to N3?}
\newline
\newline
{\color{blue}{We added the following elaboration on the benefits of N4 in section 3.4.3 (page 21):
\begin{quote}
N4 extends the popular
nonparametric nonuniform intensity normalization (N3) algorithm [63]
in two principal ways:
\begin{itemize}
  \item We replace the least squares B-spline fitting with a parallelizable alternative 
  (which we also made publicly available in the Insight Toolkit)---
  the advantages being that 1) computation is much faster 
  and 2) smoothing is not susceptible to outliers as is characteristic with
  standard least squares fitting algorithms.
  \item The fitting algorithm permits a multi-resolution approach so whereas standard 
  N3 practice is to select a single resolution at which bias correction occurs, the N4
  framework permits a multi-resolution correction where a base resolution is chosen and
  correction can then occur at multiple resolution levels each resolution being twice 
  the resolution of the previous level.
\end{itemize}  
Specifically, with respect to segmentation, there exists a third advantage with N4 over
N3 in that the former permits the specification of a probabilistic mask as opposed 
to a binary mask.  Recent demonstrations suggest improved
white matter segmentation produces better gain field estimates using
N3 [64].  Thus, when performing 3-tissue segmentation, we
may opt to use, for instance, the posterior probability map of white matter at the current
iteration as a weighted mask for input to N4.
\end{quote} }}

### 2011-02-07 Documentation/atropos/atropos.tex
Added some stuff.
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@767 fd2da0a4-d968-48df-9cb5-d8c496287cbf

Although many optimization techniques exist \citep[see the
introduction in][for a concise summary of the myriad optimization
possibilities]{Marroquin2002}---each with their characteristic

@@

K-means algorithm with only the number of classes as a prior.

@@

filename in the \texttt{--icm [1]} set of options.  An example is given in 
Figure \ref{fig:r16icm}.  For each iteration, a random ordering of the

@@

a maximum number of these ``ICM iterations'' on the command line.}}

@@

An overview of Atropos components can be gleaned, in part, from the flowchart depicted in 
Fig.~\ref{fig:flowchart}.  To provide a more intuitive interface without the overhead costs of a graphical user interface, a set of unique command line parsing classes were developed which can also provide insight to the functionality of Atropos.  
The short version of the command line help menu is given in Listing
\ref{listing:command} which is invoked by typing `{\ttfamily Atropos
  -h}' at the command prompt.  Both short and long option flags are
available and each option has its own set of possible values and
parameters introduced in a more formal way in both the previous discussion and related papers cited in the introduction.  Here we describe these options from the unique perspective of  implementation.

>>>>

Although many optimization techniques exist (see the introduction in 
\cite{Marroquin2002} for a concise summary of the myriad optimization
possibilities)---each with their characteristic

@@

K-means or Otsu thresholding algorithms with only the number of classes 
specified by the user.

@@

filename in the \texttt{--icm [1]} set of options.  An example adult brain
slice and the associated code image is given in 
Figure \ref{fig:r16icm} for an MRF neighborhood of $5 \times 5$ pixels.
This produces a maximum code label of `13'.
For each iteration, a random ordering of the

@@

a maximum number of these ``ICM iterations'' on the command line.  For
our example in Figure \ref{fig:r16icm}, this means that for each
ICM iteration, we would iterate through the image 13 times only updating
those segmentation labels associated with the current ICM code.}}

@@

An overview of Atropos components can be gleaned, in part, from the flowchart depicted in 
Fig.~\ref{fig:flowchart}.  {\color{blue}{Given a set of input images and a mask images, each is preprocessed using N4 to correct for intensity inhomogeneity.  For our brain processing pipeline, the mask is usually obtained from the standard skull-stripping preprocessing step which also uses Atropos.  Initialization can be performed in various ways using standard clustering techniques, such as K-means, to prior-based images.  This initialization is used to provide the initial estimate of the parameters of the likelihood model for each class.  These likelihoods combine with the current labeling to generate the current estimate of the posterior probabilities at each voxel for each class.  At each iteration, one can also integrate N4 by using the current posterior probability estimation of the white matter to update the estimate of bias field. }}

To provide a more intuitive interface without the overhead costs of a graphical user interface, a set of unique command line parsing classes were developed which can also provide insight to the functionality of Atropos.  
The short version of the command line help menu is given in Listing
\ref{listing:command} which is invoked by typing `{\ttfamily Atropos
  -h}' at the command prompt.  Both short and long option flags are
available and each option has its own set of possible values and
parameters introduced in a more formal way in both the previous discussion and related papers cited in the introduction.  Here we describe these options from the unique perspective of  implementation.
